---
title: "DOOR Examples"
author: "Diane Uschner"
date: "8/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Loading the package

```{r message = FALSE}
library(DOOR)
```


## Example 1

The first example data set has 26 observations. This data is from the \texttt{sample-data.csv} data set that is also included in the \texttt{./dev/} folder or on the BSC confluence page. The first few observations look like this. 

```{r}
head(sample_data)
```

First we obtain the DOOR summary for this data set:

```{r}
res <- get_door_summary(data = sample_data, tx = "seq", outcome = "DOOR" )
res
```

We use the summary to estimate the DOOR probability:

```{r}
xi <- get_door_probability(res, tx = c("A", "B")) # Probability that group A is better than group B
```

Next we compute the DOOR confidence intervals for this example using the Bootstrap, Halperin and Multinomial methods.

```{r}
get_door_prob_CI(res, tx = c("A", "B"), method = "bootstrap")
get_door_prob_CI(res, tx = c("A", "B"), method = "halperin")
get_door_prob_CI(res, tx = c("A", "B"), method = "multinom")
```


## Example 2

We now repeat the same steps for a larger dataset. This second example data set has 300 observations. This data is from the \texttt{sample_data_N_300.csv} data set that is also included in the \texttt{./dev/} folder or on the BSC confluence page. The first few observations look like this. 

```{r}
head(sample_data_N300)
```

First we obtain the DOOR summary for this data set:

```{r}
res <- get_door_summary(data = sample_data_N300, tx = "seq", outcome = "DOOR" )
res
```

We use the summary to estimate the DOOR probability:

```{r}
xi <- get_door_probability(res, tx = c("A", "B")) # Probability that group A is better than group B
```

Next we compute the DOOR confidence intervals for this example using the Bootstrap, Halperin and Multinomial methods.

```{r}
get_door_prob_CI(res, tx = c("A", "B"), method = "bootstrap")
get_door_prob_CI(res, tx = c("A", "B"), method = "halperin")
get_door_prob_CI(res, tx = c("A", "B"), method = "multinom")
```



