---
title: "Desirability of Outcome Ranking"
author: Michelle Earley, Weixiao Dai, Qing Pan, Toshi Hamasaki, Scott Evans, Diane
  Uschner
date: "February 5, 2020"
documentclass: jss
classoption: article
output: 
  pdf_document:
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

# Methods
Consider the situation of a study with treatment groups $A$ and $B$, and DOOR outcome $Y$ with $K$ levels. 
Let the total sample size be denoted by $n$, where $n = n_A + n_B$ and $n_t$ denotes
the sample size in group $t \in \{A, B\}$. Let $n_{t,k}$ denote number of subjects in group $t$ that experienced outcome $Y = k$.
The situation is presented in Table \ref{tab:door-summary}.

\begin{table}[ht]
\centering
\begin{tabular}{ccc}
\toprule
Outcome & \multicolumn{2}{c}{Treamtent Group} \\
$Y = k$ & $A$ & $B$\\
\midrule
1 & $n_{A,1}$ & $n_{B,1}$\\
2 & $n_{A,2}$ & $n_{B,2}$\\
\vdots & \vdots & \vdots \\
$K$ & $n_{A,K}$ & $n_{B,K}$\\
\midrule
Total & $n_A$ & $n_B$\\
\bottomrule
\end{tabular}
\caption{DOOR summary table of sample sizes by DOOR outcome and treatment group.}
\label{tab:door-summary}
\end{table}

Denote the vector of sample sizes in group $t \in \{A, B\}$ by outcome as $v_t = (n_{t,1}, \dots, n_{t,K})^T$.

## Point estimates 

### The DOOR probability

The DOOR probability is defined as the probability $p = P(Y_A \geq Y_B)$ that a randomly selected participant
in group $A$ experiences a better outcome than a randomly selected participant in group $B$. The DOOR probability
$p$ can be estimated by the fractions of tuples $(Y_{A,i}, Y_{B,j})$, where $Y_{A,i} \geq Y_{B,j}$, namely

\begin{equation}
\hat{p} = \sum_{i = 1}^{n_A}\sum_{j = 1}^{n_B} \frac{1(Y_{A,i} \geq Y_{B,j})}{n_A\cdot n_B}.
\end{equation}

We can use the DOOR summary in table \ref{tab:door-summary} to calculate this estimate more efficiently. We recognize 
that for each participant in group $B$ that had outcome $Y = k$, there are $n_{A,k}$ participants in group $A$ that had 
the same outcome, and $\sum_{j = k+1}^K n_{A, j}$ participants that had a larger outcome. Denote by $M$ the symmetric matrix
that contains the shifted sample size vector $v_A$ of group $A$ in rows and columns:

\begin{equation}
M = 
\begin{pmatrix}
n_{A,2} & n_{A,3} & \dots & n_{A,K} & 0\\
n_{A,3} & n_{A,4} & \dots & 0 & 0\\
\vdots & \vdots  & \dots & \vdots  & \vdots \\
n_{A,K} & 0  & \dots & 0 & 0\\
0  & 0 & \dots & 0 & 0\\
\end{pmatrix}
\end{equation}.

We can then estimate the DOOR probability by taking the matrix product of $M$ and $v_B$ for the probability of observing a larger outcome in group $A$ than in $B$, and adding half the scalar product of $v_A$ and $v_B$ for the probability of observing the same outcome in both groups, and dividing by the total number of possible comparisons $n_A\cdot n_B$ 

\begin{equation}
\hat{p} = \frac{1}{n_A \cdot n_B}\cdot(1^{T} M v_B + \frac{1}{2} v_A^{T} v_B.)
\end{equation}

### Difference in means

### Partial Score


## Conference Intervals
  
### Bootstrap CI for the DOOR probability

### Wilcoxon rank sum method

### Asymptotic Formula for the DOOR CI

## Implementation

We created an \proglang{R} package for the DOOR probability. The fpllowing code generates Table \ref{tab:door-summary} for an example of
$n = 300$ participants

```{r message = FALSE}
library(DOOR)

set.seed(123)
n <- 300
seq <- rep(c("A", "B"), each = n/2)
DOOR <- sample(1:4, size = n, replace = TRUE)
data <- data.frame(seq, DOOR)
(res <- get_door_summary(data, "seq", "DOOR"))
```

The following command creates Figure \ref{fig:door-comparison} that visualizes Table \ref{tab:door-summary}.

```{r, fig.cap="\\label{fig:door-comparison}Visualization of the DOOR summary"}
plot_door_comparison(res)
```



