---
title: "The Halperin Method"
author: "Diane Uschner"
date: "8/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

# Method

We calculate the single steps of the Halperin algorithm for the first of the simulation settings of the Halperin et al. (1989) paper.

## Settings

```{r}
m <- n <- 250
K <- 3

p1 <- c(1/3, 1/3, 1/3)
p2 <- c(1/3, 1/3, 1/3)

q1 <- 1 - p1
q2 <- 1 - p2
```

## Intermediate steps of the algorithm - Maximum Likelihood Estimators

Cumulative sums of the probability vector:

```{r}
(P2 <- apply(sapply(1:K, function(k) lead(p2, n = k, default = 0)), 2, sum))
(P1 <- apply(sapply(1:K, function(k) dplyr::lag(p1, n = K-k+1, default = 0)), 2, sum))
```

The DOOR probability for treatment 1 better than treatment 2

```{r}
(xi <- sum(p1 %*% P2) + 0.5 * sum(p1*p2))
```
The coefficients 

```{r}
(A <- sum((p1*(P2 + (p2/2))^2)[-K]) + 0.25 * p1[K]*(p2[K])^2)
(B <- sum((p2*(P1 + (p1/2))^2)[-1]) + 0.25 * p2[1]*(p1[1])^2)
```

### Variance - Maximum Likelihood estimator - Formula 3.2

```{r}
(V1 <- 1/(m*n)*(xi - (m + n -1) * xi^2 + (n-1) * A + (m-1)*B - 0.25* sum(p1*p2)))
```


## Unbiased Estimators

### A hat hat
```{r}
(s1 <- sum(p1*(q2*P2 - P2^2)))
(corrA <- 1/(n-1) * s1 -  1/(4* (n-1)) * sum(p1*p2*q2))
(Ahh <- A - corrA)
```

### B hat hat

```{r}
(s2 <- sum(p2 * (q1 * P1 - P1^2)))
(corrB <- 1/(m-1) * s2 -  1/(4* (m-1)) * sum(p2*p1*q1))
(Bhh <- B - corrB)
```

### Formula 3.9 - Unbiased estimator of xi* (1 - xi)

```{r}
(xi2 <- ((m*n - m - n +2) * xi - m*n*xi^2)/((m-1)*(n-1)) + Ahh/(m-1) + Bhh/(n-1))
```


```{r}
# Formula 3.6
theta <- ((m+n-2) * xi - (n-1)*Ahh - (m-1) * Bhh)/((m + n - 2)* xi2)
(theta <- max(min(theta, 1), 0))
```


### Variance - unbiased estimator - Formula 3.5, with 3.6 - 3.9 substituted.
```{r}
(V2 <- 1/(m*n)*(m + n - 1 - (m + n -2) * theta) * xi2)
```

## Confidence Interval

```{r}
(gamma <- (m + n - 1) - (m + n - 2) * theta)
(C <- gamma * qchisq(1-0.05, 1)/(m*n))
(b <- sqrt(C^2 + 4 * C * xi*(1-xi)))

(UL <- (C + 2 * xi + b)/(2 * (C+1)))
(LL <- (C + 2 * xi - b)/(2 * (C+1)))

```



# Applying the method to simulated data

Below a comparison of the halperin method for simulated data

```{r message=FALSE}
library(DOOR)

set.seed(1986)
N <- 200
seq <- rep(c("A", "B"), each = N/2)
DOOR <- sample(1:3, size = N, replace = TRUE)
data <- data.frame(seq, DOOR)
(res <- get_door_summary(data, "seq", "DOOR"))
get_door_probability(res)
get_door_prob_CI(res, tx = c("A", "B"), method = "halperin")
get_door_prob_CI(res, tx = c("A", "B"), method = "bootstrap")
get_door_prob_CI(res, tx = c("A", "B"), method = "multinom")
```
